Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-10-03 paper_comparison/paper_comparison/paper_comparison.Rmd
fill in analysis

As discussed in @majumder_validation_2013, individual's skill will affect the number of observers who identify the actual data plot from the lineup. Thus, the power of a visual test depends on the subject-specific abilities. Previously, it was addressed by modelling the probability of a subject $i$ correctly picking the actual data plot from a lineup $l$ using a mixed-effect logistic regression with the subject being treated as a random effect [@majumder_validation_2013]. However, having this probability is insufficient to determine the power of a visual test allowed for multiple selections as it doesn't provide information about the number of selections made by the subject for p-value calculation. Instead, we directly estimated the probability of a lineup being rejected using a logistic regression with the natural logarithm of the effect size as the only regressor. The same model was applied to different subsets of the survey data for making comparison of the power.

@@

## Results

### Power comparison

```{r}
vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                pivot_longer(F_p_value:SW_p_value), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  facet_wrap(~x_dist) +
  ggtitle("poly")
```

### Data overview

@@

<!-- ## Nonlinearity -->

<!-- -   Power curve overall -->
<!-- -   Four lineups also shown, selected from ones close to the visual power curve from the uniform treatment -->
<!-- -   Compare visual power against different factors -->

<!-- ## Heteroskedasticity -->

<!-- -   Power curve overall -->
<!-- -   Four lineups also shown, selected from ones close to the visual power curve from the uniform treatment -->
<!-- -   Compare visual power against different factors -->

<!-- ## Demographic summary -->

<!-- Table \ref{tab:demographic-table} tabulates the number of subjects, preferred pronouns, education backgrounds, age groups, and previous experience in visual experiment. Figure \ref{fig:demographic-summary-plot} visualizes the marginal distribution of each of the category. The collated data was a balanced sample among male and female. Most of the ages of subject were between 18 to 39, while many of them were between 18 to 24. Very few subjects were awarded degree higher than Bachelor Degree. Around 40% of subjects had previous experience in visual experiment. -->

<!-- ```{r results="asis"} -->

<!-- polynomials %>% -->
<!--   mutate(exp = 1) %>% -->
<!--   bind_rows(mutate(heter, exp = 2)) %>% -->
<!--   group_by(exp, set) %>% -->
<!--   slice_sample(n = 1) %>% -->
<!--   ungroup() %>% -->
<!--   count(pronoun, education, age_group, previous_experience) %>% -->
<!--   arrange(desc(n)) %>% -->
<!--   knitr::kable(col.names = c("Preferred pronoun", "Education", "Age group", "Previous experience", "Subject"), -->
<!--                escape = FALSE, -->
<!--                format = "latex", -->
<!--                linesep = "", -->
<!--                booktabs = TRUE,  -->
<!--                caption  = 'Summary of demographic information',  -->
<!--                label = "demographic-table" ) %>% -->
<!--   cat() -->
<!-- ``` -->

<!-- ```{r demographic-summary-plot, fig.cap="Summary of demographic information. \\label{fig:demographic-summary-plot}"} -->
<!-- polynomials %>% -->
<!--   mutate(exp = 1) %>% -->
<!--   bind_rows(mutate(heter, exp = 2)) %>% -->
<!--   group_by(exp, set) %>% -->
<!--   summarise(across(age_group:previous_experience, first)) %>% -->
<!--   mutate(education = fct_relevel(education, "High School or below")) %>% -->
<!--   ggplot() + -->
<!--   geom_bar(aes(y = education), orientation = "y") -> education_plot -->

<!-- polynomials %>% -->
<!--   mutate(exp = 1) %>% -->
<!--   bind_rows(mutate(heter, exp = 2)) %>% -->
<!--   group_by(exp, set) %>% -->
<!--   summarise(across(age_group:previous_experience, first)) %>% -->
<!--   ggplot() + -->
<!--   geom_bar(aes(age_group)) -> age_plot -->

<!-- polynomials %>% -->
<!--   mutate(exp = 1) %>% -->
<!--   bind_rows(mutate(heter, exp = 2)) %>% -->
<!--   group_by(exp, set) %>% -->
<!--   summarise(across(age_group:previous_experience, first)) %>% -->
<!--   mutate(pronoun = fct_relevel(pronoun, "He", "She", "They")) %>% -->
<!--   ggplot() + -->
<!--   geom_bar(aes(pronoun)) -> gender_plot -->

<!-- polynomials %>% -->
<!--   mutate(exp = 1) %>% -->
<!--   bind_rows(mutate(heter, exp = 2)) %>% -->
<!--   group_by(exp, set) %>% -->
<!--   summarise(across(age_group:previous_experience, first)) %>% -->
<!--   ggplot() + -->
<!--   geom_bar(aes(previous_experience)) -> experience_plot -->

<!-- patchwork::wrap_plots(education_plot, age_plot, gender_plot, experience_plot) -->
<!-- ``` -->

<!-- ## Model fitting -->

<!-- For each parameter combination, effect $E$ is derived from the Kullback-Leibler divergence (see [appendix ref here]) formulated as: -->

<!-- ```{=tex} -->
<!-- \begin{equation} \label{eq:effect-size-ex1} -->
<!-- E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b, -->
<!-- \end{equation} -->
<!-- ``` -->
<!-- where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$. -->

<!-- The logistic regression is fit using $log_e(E)$ as the only fixed effect covariate for the power of visual test formulated as: -->

<!-- ```{=tex} -->
<!-- \begin{equation} \label{eq:logistic-regression-1-1} -->
<!-- Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})), -->
<!-- \end{equation} -->
<!-- ``` -->
<!-- where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$. -->

<!-- To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors. This includes [expansion]. -->

<!-- Table [table ref here] shows the parameter estimates of the logistic regressions. [Discussion about the numeric estimates here] -->

>>>>

As discussed in @majumder_validation_2013, individual's skill will affect the number of observers who identify the actual data plot from the lineup. Thus, the power of a visual test depends on the subject-specific abilities. Previously, it was addressed by modelling the probability of a subject $i$ correctly picking the actual data plot from a lineup $l$ using a mixed-effect logistic regression with the subject being treated as a random effect [@majumder_validation_2013]. However, having this probability is insufficient to determine the power of a visual test allowed for multiple selections as it doesn't provide information about the number of selections made by the subject for p-value calculation.

Instead, we directly estimated the probability of a lineup being rejected using a logistic regression with the natural logarithm of the effect size as the only regressor formulated as:

```{=tex}
\begin{equation} \label{eq:logistic-regression-1-1}
Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})),
\end{equation}
```
where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.

Effect $E$ is derived from the Kullback-Leibler divergence (see [appendix ref here]) formulated as:

```{=tex}
\begin{equation} \label{eq:effect-size-ex1}
E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b,
\end{equation}
```
where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.

To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors. This includes [expansion].

@@

# Results

## Data overview

@@

## Power comparison

Figure \ref{fig:power-poly-uniform} shows the estimated power of visual test against natural logarithm of the effect with comparison to the power of an exact test - F-test, and the power of two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions. In overall, the power of all four tests increases as the effect becomes larger. The power curve of F-test climbs aggressively from 25% to around 90% as $log_e(E)$ increases from 0 to 2, while others respond inactively to the change of effect and remain lower than 25% throughout the period, showing that as an exact test, the F-test is relatively more sensitive to the type of model defects that being considered. The power of visual test arises steadily and nearly linearly to around 90% as $log_e(E)$ increases from 2 to 5, suggesting that the effect starts to make noticeable impact on the degree of the presence of the designed visual features. Other two inappropriate conventional tests shows improvement at the same time but at a lower rate. This coincides the point made by @cook1982residuals mentioned in \ref{hypothesis-testing} that residual-based tests for a specific type of model defect are sensitive to other types of model defects. At $log_e(E) = 6$, the power curve of F-test reaches almost 100% followed by the visual test by a small margin. The power of Breusch--Pagan test and Shapiro--Wilk test reach around 75% and 30% respectively.

What truly impress us is the huge difference between the estimated power of visual test and the estimated power of F-test. The margin is largest at around $log_e(E) = 2$. An example lineup is included in Figure \ref{fig:power-overview} where none of subjects detect the actual data plot positioned at panel 14. It demonstrates that at this level of difficulty, the designed visual feature is rarely visible, making the actual data plot indistinguishable from residual plots simulated from the assumed model. From a communication perspective, given the fact that the visual difference is unperceivable, the argument that non-linearity present in the fitted model is less convincing to the public even though it is true. At around $log_e(E) = 3$, the margin gets smaller as the chance of identifying the actual data plot becomes larger. At this level of difficulty, the designed visual features are usually detectable but it may not stand out from the lineup as other null plots may happen to include outliers or visual patterns that are are considered to be more attractive by human, and thus recognized as the most different plot. Without knowing the designed visual features beforehand, it is actually hard to identify the actual data plot by pure image comparison. The corresponding example lineup for $log_e(E) = 3$ shown in Figure \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at panel 20, where two out of five subjects detect it. It can be observed that a M-shape is presented in plot 20, but the signal is not strong enough to attract all five subjects, resulting in a visual p-value sightly above the desired significance level $\alpha = 0.05$. At $log_e(E) = 4$ and $log_e(E) = 6$, the designed visual features become much clear and attractive, leading to a high percentage of rejection of the null hypothesis. Figure \ref{fig:fig:power-poly-uniform} gives example lineups of such cases. It is also interesting to observe that the power curve of visual test behaves similarly to those inappropriate conventional tests until the designed visual features really becomes visible.

```{r fig:power-poly-uniform, fig.cap = "..."}
vi_survey_processed %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) + 
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                filter(x_dist == "uniform") %>%
                pivot_longer(F_p_value:SW_p_value), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  theme_light(base_size = 5) -> p

VI_MODEL$plot_lineup(vi_lineup[["poly_39"]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(base_size = 5),
                     remove_axis = TRUE)

VI_MODEL$plot_lineup(vi_lineup[["poly_415"]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(base_size = 5),
                     remove_axis = TRUE) -> p1

VI_MODEL$plot_lineup(vi_lineup[["poly_131"]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(base_size = 5),
                     remove_axis = TRUE) -> p2

VI_MODEL$plot_lineup(vi_lineup[["poly_322"]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(base_size = 5),
                     remove_axis = TRUE) -> p3

VI_MODEL$plot_lineup(vi_lineup[["poly_378"]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(base_size = 5),
                     remove_axis = TRUE) -> p4

design <- c(patchwork::area(1, 1, 1, 8),
            patchwork::area(2, 1, 2, 4), 
            patchwork::area(2, 5, 2, 8), 
            patchwork::area(3, 1, 3, 4), 
            patchwork::area(3, 5, 3, 8))

patchwork::wrap_plots(p, B = p1, C = p2, D = p3, E = p4, design = design)
```

## Distribution of regressor

The impact of the distribution of $X_raw$ on the power is shown in Figure \ref{fig:dist-power}. The power curve of F-test is stable across different distributions, while the visual test has a steeper power curve for normal and uniform distribution. BP-test performs worse for discrete uniform distribution and uniform distribution but has relatively high power for normal distribution. SW-test outperforms BP-test for discrete uniform distribution but remains as the worst test for other distributions. The results indicate those inappropriate residual-based tests are sensitive to the distribution of the regressor.

```{r dist-power, fig.cap="..."}
vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) +
  stat_smooth(geom = "line",
              data = ~boot_y(.x, reject_dirichlet, times = 500),
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"),
              method = "glm",
              method.args = list(family = binomial),
              se = FALSE,
              alpha = 0.03) +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                pivot_longer(F_p_value:SW_p_value),
              aes(log_effect_size, as.numeric(value < 0.05), col = name),
              method = "glm",
              method.args = list(family = binomial),
              se = FALSE) +
  facet_wrap(~x_dist) +
  theme_light(base_size = 5)
```

### 2022-10-03 paper_comparison/paper_comparison/paper_comparison.tex
fill in analysis

As discussed in \citet{majumder_validation_2013}, individual's skill
will affect the number of observers who identify the actual data plot
from the lineup. Thus, the power of a visual test depends on the
subject-specific abilities. Previously, it was addressed by modelling
the probability of a subject \(i\) correctly picking the actual data
plot from a lineup \(l\) using a mixed-effect logistic regression with
the subject being treated as a random effect
\citep{majumder_validation_2013}. However, having this probability is
insufficient to determine the power of a visual test allowed for
multiple selections as it doesn't provide information about the number
of selections made by the subject for p-value calculation. Instead, we
directly estimated the probability of a lineup being rejected using a
logistic regression with the natural logarithm of the effect size as the
only regressor. The same model was applied to different subsets of the
survey data for making comparison of the power.

@@

\hypertarget{results}{%
\subsection{Results}\label{results}}

\hypertarget{power-comparison}{%
\subsubsection{Power comparison}\label{power-comparison}}

\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-7-1}

\hypertarget{data-overview}{%
\subsubsection{Data overview}\label{data-overview}}

@@

>>>>

As discussed in \citet{majumder_validation_2013}, individual's skill
will affect the number of observers who identify the actual data plot
from the lineup. Thus, the power of a visual test depends on the
subject-specific abilities. Previously, it was addressed by modelling
the probability of a subject \(i\) correctly picking the actual data
plot from a lineup \(l\) using a mixed-effect logistic regression with
the subject being treated as a random effect
\citep{majumder_validation_2013}. However, having this probability is
insufficient to determine the power of a visual test allowed for
multiple selections as it doesn't provide information about the number
of selections made by the subject for p-value calculation.

Instead, we directly estimated the probability of a lineup being
rejected using a logistic regression with the natural logarithm of the
effect size as the only regressor formulated as:

\begin{equation} \label{eq:logistic-regression-1-1}
Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})),
\end{equation}

where \(\Lambda(.)\) is the standard logistic function given as
\(\Lambda(z) = exp(z)/(1+exp(z))\).

Effect \(E\) is derived from the Kullback-Leibler divergence (see
{[}appendix ref here{]}) formulated as:

\begin{equation} \label{eq:effect-size-ex1}
E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b,
\end{equation}

where \(diag(.)\) is the diagonal matrix constructed from the diagonal
elements of \(\boldsymbol{R}_a\).

To study various factors contributing to the power of the visual test,
the same logistic regression model is fit on different subsets of the
collated data grouped by levels of factors. This includes
{[}expansion{]}.

@@

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{data-overview}{%
\subsection{Data overview}\label{data-overview}}

@@

\hypertarget{power-comparison}{%
\subsection{Power comparison}\label{power-comparison}}

Figure \ref{fig:power-poly-uniform} shows the estimated power of visual
test against natural logarithm of the effect with comparison to the
power of an exact test - F-test, and the power of two other
residual-based conventional tests commonly used in regression
diagnostics but for testing other departures from the model assumptions.
In overall, the power of all four tests increases as the effect becomes
larger. The power curve of F-test climbs aggressively from 25\% to
around 90\% as \(log_e(E)\) increases from 0 to 2, while others respond
inactively to the change of effect and remain lower than 25\% throughout
the period, showing that as an exact test, the F-test is relatively more
sensitive to the type of model defects that being considered. The power
of visual test arises steadily and nearly linearly to around 90\% as
\(log_e(E)\) increases from 2 to 5, suggesting that the effect starts to
make noticeable impact on the degree of the presence of the designed
visual features. Other two inappropriate conventional tests shows
improvement at the same time but at a lower rate. This coincides the
point made by \citet{cook1982residuals} mentioned in
\ref{hypothesis-testing} that residual-based tests for a specific type
of model defect are sensitive to other types of model defects. At
\(log_e(E) = 6\), the power curve of F-test reaches almost 100\%
followed by the visual test by a small margin. The power of
Breusch--Pagan test and Shapiro--Wilk test reach around 75\% and 30\%
respectively.

What truly impress us is the huge difference between the estimated power
of visual test and the estimated power of F-test. The margin is largest
at around \(log_e(E) = 2\). An example lineup is included in Figure
\ref{fig:power-overview} where none of subjects detect the actual data
plot positioned at panel 14. It demonstrates that at this level of
difficulty, the designed visual feature is rarely visible, making the
actual data plot indistinguishable from residual plots simulated from
the assumed model. From a communication perspective, given the fact that
the visual difference is unperceivable, the argument that non-linearity
present in the fitted model is less convincing to the public even though
it is true. At around \(log_e(E) = 3\), the margin gets smaller as the
chance of identifying the actual data plot becomes larger. At this level
of difficulty, the designed visual features are usually detectable but
it may not stand out from the lineup as other null plots may happen to
include outliers or visual patterns that are are considered to be more
attractive by human, and thus recognized as the most different plot.
Without knowing the designed visual features beforehand, it is actually
hard to identify the actual data plot by pure image comparison. The
corresponding example lineup for \(log_e(E) = 3\) shown in Figure
\ref{fig:fig:power-poly-uniform} has the actual data plot positioned at
panel 20, where two out of five subjects detect it. It can be observed
that a M-shape is presented in plot 20, but the signal is not strong
enough to attract all five subjects, resulting in a visual p-value
sightly above the desired significance level \(\alpha = 0.05\). At
\(log_e(E) = 4\) and \(log_e(E) = 6\), the designed visual features
become much clear and attractive, leading to a high percentage of
rejection of the null hypothesis. Figure
\ref{fig:fig:power-poly-uniform} gives example lineups of such cases. It
is also interesting to observe that the power curve of visual test
behaves similarly to those inappropriate conventional tests until the
designed visual features really becomes visible.

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/fig:power-poly-uniform-1} \caption{...}\label{fig:fig:power-poly-uniform-1}
\end{figure}
\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/fig:power-poly-uniform-2} \caption{...}\label{fig:fig:power-poly-uniform-2}
\end{figure}

\hypertarget{distribution-of-regressor}{%
\subsection{Distribution of regressor}\label{distribution-of-regressor}}

The impact of the distribution of \(X_raw\) on the power is shown in
Figure \ref{fig:dist-power}. The power curve of F-test is stable across
different distributions, while the visual test has a steeper power curve
for normal and uniform distribution. BP-test performs worse for discrete
uniform distribution and uniform distribution but has relatively high
power for normal distribution. SW-test outperforms BP-test for discrete
uniform distribution but remains as the worst test for other
distributions. The results indicate those inappropriate residual-based
tests are sensitive to the distribution of the regressor.

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/dist-power-1} \caption{...}\label{fig:dist-power}
\end{figure}
